Switch equivalence and weak prime domains for left-linear DPO rewriting systems

### 2025-04-17 full_version/leftlinear.tex
switch

		\node[font=\scriptsize, below] at (G1.south) {$G_0$};
		\node[font=\scriptsize, below] at (D1.south) {$D_0$};      
		\node[font=\scriptsize, below] at (G2.south) {$G_1$};
		\node[font=\scriptsize, below] at (D2.south) {$D_1'$};      
		\node[font=\scriptsize, below] at (G3.south) {$G_2'$};
		\node[font=\scriptsize, below] at (D3.south) {$D_2'$};      
		\node[font=\scriptsize, below] at (G4.south) {$G_3$};

@@

\begin{lemma}\label{lem:abs}
	Let $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be the following derivation:

	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]
	and suppose that the derivation $\der{E}$ below is a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

	Then the following hold true:
	\begin{enumerate}
		\item if $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ is an abstraction equivalence between $\der{D}$ and another derivation $\der{F}=\{\dder{F}_i\}_{i=0}^1$, then $\phi^{-1}_{G_0}*\der{E}*\phi_{G_2}$ is a pre-switch for $\der{F}$ along $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$;
		\item if $\{\psi_X\}_{X\in \Deltamin(\der{E})}$ is an abstraction equivalence between $\der{E}$ and $\der{G}=\{\dder{G}_{i}\}_{i=0}^1$, then $\psi_{G_0}*\der{G}*\psi^{-1}_{G_2}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$.
	\end{enumerate}
	\begin{remark}Recall that, by \Cref{rem:seqabs}, $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$ is an independence pair between $\dder{F}_0$ and $\dder{F}_1$.
	\end{remark}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item To prove this, it is enough to compute:
		\[\begin{split}
			m_{\der{F},0}=&\phi_{G_0}\circ m_0\\=&\phi_{G_0}\circ f'_0\circ i'_1\\\phi_{G_0}\circ	m'_0=&\phi_{G_0}\circ f_0\circ i_1\\=&f_{\der{F},0}\circ \phi_{D_{0}}\circ i_1
		\end{split} \quad \begin{split}
	h_{\der{F},1}=&\phi_{G_2}\circ h_1\\=&\phi_{G_2}\circ g_1' \circ i_0' \\	\phi_{G_2}\circ h'_1=&\phi_{G_2}\circ g_1\circ i_0\\=&g_{\der{F}, 1}\circ \phi_{D_1}\circ i_0\end{split} 
	\]

		\item

		And we can conclude. \qedhere 
	\end{enumerate}
\end{proof}

@@

	Let $\X$ be the  
	category associated with this order, which by \Cref{rem:iso} is
	$\mathsf{I(\X)}$-adhesive. Consider a rewriting
	system whose set of rules contains the following 
	\[\xymatrix{a & a \ar[r]^{\sqsubseteq} \ar@{>->}[l]_{\id{a}} & 0 & a & a
		\ar[r]^{\sqsubseteq} \ar@{>->}[l]_{\id{a}} & b }\]
	We can then consider the  following derivation
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$.
	 \[
   \xymatrix@C=25pt{a \ar[d]_{(a,c)}&& \ar@{>->}[d]_{k_0}\ar[ll]_{\id{a}}
		    \ar[r]^{(a,0)} & 0 \ar@/^.35cm/[drrr]|(.315)\hole_(.45){\id{0}}
		    \ar[dr]|(.3)\hole_{\id{0}} && a \ar@/_.35cm/[dlll]^(.45){\hspace{5pt}(a,c)}
		     \ar[dl]|(.3)\hole^{(a,0)}& a \ar[d]^{(a,0)}\ar@{>->}[l]_{\id{a}}
		     \ar[rr]^{(a,b)} && b \ar[d]^{(b,0)}\\c &&
		     \ar@{>->}[ll]^{\id{c}} c \ar[rr]_{(c,0)}&& 0 &&
		     \ar@{>->}[ll]^{\id{0}} 0 \ar[rr]_{\id{0}}&& 0}\]

@@

\begin{definition}[Switch] 	Let $(\X, \R)$ be a left-linear rewriting system, with $\X$ an $\mathcal{M}$-adhesive category.
	Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the diagram below, where the cone with vertex $P$ is a pullback and $u_0\colon K_0\to P$ and $u_1\colon K_1\to P$ are the arrows built in \Cref{prop:tec}.
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0 \ar @/^.35cm/[ddrr]_(.2){u_0}
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}|(.68)\hole
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}|(.68)\hole 
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar @/_.35cm/[ddll]^(.2){u_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}|(.7)\hole&& G_1 && \ar@{>->}[ll]^{f_1}|(.7)\hole D_1
		\ar[rr]_{g_1}&& G_2 \\ & & & & P \ar[urr]_{p_1} \ar@{>->}[ull]^{p_0}}
	\]

	Let $\der{E}=\{\dder{E}_i\}_{i=0}^1$, depicted below, be a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

@@

\begin{lemma}
Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation of length $2$ and $(i_0, i_1)$ an independence pair between $\dder{D}_0$ and $\dder{D}_1$. Given a pre-switch $\der{E}=\{\dder{E}_{i}\}_{i=0}^1$ for $\der{D}$ along $(i_0, i_1)$, the following are equivalent:
	\begin{enumerate}
		\item $\der{E}$ is a switch;
		\item if $(P, \{p_i\}_{i=0}^1)$ is a pullback for $f_{\der{D},1}$ along $g_{\der{D},0}$, then $\iota_{\der{E}, G_2} \circ g_1\circ p_1=\iota_{\der{E}, G'_1}\circ g'_0\circ z_0$.
	\end{enumerate}
\end{lemma}
\begin{proof}$(1 \Rightarrow 2)$. If $\der{E}$ is a switch then we have
	\begin{align*}
		\iota_{\der{E}, G_2} \circ g_1\circ p_1&=\iota_{\der{E}, G_2} \circ g'_1\circ z_1\\&=\iota_{\der{E}, D'_1}\circ z_1\\&=\iota_{\der{E}, G'_1}\circ f'_1\circ z_1\\&=\iota_{\der{E}, G'_1}\circ g'_0\circ z_0
	\end{align*}

@@

	By \Cref{rem:zero1} the inner square is a pullback, proving that the dotted arrow $z_1\colon P\to D'_1$ exists.  Since, by \Cref{lem:colim} $\iota'_{1, G'_1}$ is in $\mathcal{M}$, we conclude that $f'\circ z_1=g'\circ z_0$ as wanted.
\end{proof}

@@

\begin{proposition}\label{prop:pretosw} 	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose
	that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$ and
	$\der{F}=\{\dder{F}_i\}_{i=0}^1$ are pre-switches of $\dder{D}_0$ and
	$\dder{D}_1$ along the same independence pair $(i_0, i_1)$. Let $\{\phi_X\}_{X\in \Deltamin(\der{E})}$ be the abstraction equivalence built in \Cref{thm:switch_uni}. If $\der{F}$ is a switch then also $\der{E}*\phi_{G_2}$ is one.
\end{proposition}
\begin{proof}
Suppose that $\der{D}$ is given by the following derivation.

	By \Cref{thm:switch_uni} we have a diagram as the one below.
	\[\xymatrix@C=22pt{G_{0} && \ar@{>->}[ll]_{f_{\der{F},0}} D_{\der{F},0} \ar[rr]^{g_{\der{F},0}}&& G_{\der{F},1} && \ar@{>->}[ll]_{f_{\der{F},1}} D_{\der{F},1} \ar[rr]^{g_{\der{F},1}}&& G_{2}\\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}&& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_{1} \ar@/^.35cm/[drrr]|(.315)\hole_(.4){j_0} \ar[dr]|(.3)\hole_{h_{\der{E},0}} \ar@/_.35cm/[urrr]|(.315)\hole^(.4){a_0} \ar[ur]|(.3)\hole^{h_{\der{F},0}}& &L_{0} \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.3)\hole^{m_{\der{E},1}} \ar@/^.35cm/[ulll]_(.4){a_1} \ar@/_.3cm/[ul]|(.252)\hole_(.4){m_{\der{F},1}}& K_{0} \ar[d]_{k_{\der{E},1}} \ar[u]^{k_{\der{F},1}}\ar@{>->}[l]_{l_{0}} \ar[rr]^{r_0} && R_{0} \ar[d]_{g_1\circ i_0} \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} && \ar[ll]^{f_{\der{E},0}} D_{\der{E},0} \ar@{>}@/^.5cm/[uu]^(.33){\phi_{D_{\der{E},0}}}|\hole \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} \ar@{>}[uu]_(.76){\phi_{G_{\der{E}, 1}}}|(.45)\hole |(.55)\hole&& \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1} \ar@{>}@/_.5cm/[uu]_(.33){\phi_{D_{\der{E},1}}}|\hole\ar[rr]_{g_{\der{E},1}}&& G_{2}\ar@{>}@/_.5cm/[uu]_{\phi_{G_2}}}\]

	Moreover, by definition $\der{E}*\phi_{G_2}$

@@

We are now ready to introduce the fundamental concept of this section.

@@

>>>>

		\node[font=\scriptsize, below] at (G1.south) {$G_0$};
		\node[font=\scriptsize, below] at (D1.south) {$D_0$};      
		\node[font=\scriptsize, below] at (G2.south) {$G_1$};
		\node[font=\scriptsize, below] at (D2.south) {$D_1'$};      
		\node[font=\scriptsize, below] at (G3.south) {$G_2'$};
		\node[font=\scriptsize, below] at (D3.south) {$D_2'$};      
		\node[font=\scriptsize, below] at (G4.south) {$G_2$};

@@

\begin{lemma}\label{lem:abs}
	Let $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be the following derivation:
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]
	and suppose that the derivation $\der{E}$ below is a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

	Then the following hold true:
	\begin{enumerate}
		\item if $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ is an abstraction equivalence between $\der{D}$ and another derivation $\der{F}=\{\dder{F}_i\}_{i=0}^1$, then $\phi^{-1}_{G_0}*\der{E}*\phi_{G_2}$ is a pre-switch for $\der{F}$ along $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$;
		\item if $\{\psi_X\}_{X\in \Deltamin(\der{E})}$ is an abstraction equivalence between $\der{E}$ and $\der{G}=\{\dder{G}_{i}\}_{i=0}^1$, then $\psi_{G_0}*\der{G}*\psi^{-1}_{G_2}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$.
	\end{enumerate}
	\begin{remark}Recall that, by \Cref{rem:seqabs}, $(\phi_{D_1}\circ i_0, \phi_{D_0}\circ i_1 )$ is an independence pair between $\dder{F}_0$ and $\dder{F}_1$.
	\end{remark}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item To prove this, it is enough to compute:
		\[\begin{split}
			m_{\der{F},0}=&\phi_{G_0}\circ m_0\\=&\phi_{G_0}\circ f'_0\circ i'_1
		\end{split} \quad \begin{split}\phi_{G_0}\circ	m'_0=&\phi_{G_0}\circ f_0\circ i_1\\=&f_{\der{F},0}\circ \phi_{D_{0}}\circ i_1
	 \end{split} 
	\]
	\[\begin{split}
	h_{\der{F},1}=&\phi_{G_2}\circ h_1\\=&\phi_{G_2}\circ g_1' \circ i_0'
	\end{split} \qquad \begin{split}
		\phi_{G_2}\circ h'_1=&\phi_{G_2}\circ g_1\circ i_0\\=&g_{\der{F}, 1}\circ \phi_{D_1}\circ i_0
	\end{split}\]

		\item By \Cref{rem:seqabs} we know that $(\psi_{D'_1}\circ i'_0,\psi_{D'_0}\circ i'_1)$ is an independence pair between $\dder{G}_0$ and $\dder{G}_1$. By \Cref{rem:absequi} and again by \Cref{rem:seqabs} we conclude that  $(\psi_{D'_1}\circ i'_0,\psi_{D'_0}\circ i'_1)$ witnesses the sequential independence of the two steps making $\psi_{G_0}*\der{G}*\psi^{-1}_{G_2}$. Moreover:

		\[\begin{split}
			m_{0}=&f'_0\circ i'_1\\=&\psi^{-1}_{G_0}\circ f_{\der{G},0}\circ \psi_{D'_0}\circ i'_1
		\end{split} \qquad \begin{split}\psi^{-1}_{G_0}\circ	m_{\der{G},0}=&m'_0\\=&f_0\circ i_1 \end{split} \]
		\[\begin{split}h_1=&g'_1\circ i'_0\\=&\psi^{-1}_{G_2}\circ g_{\der{G},1} \circ\psi_{D'_0}\circ  i_0'
		\end{split} \qquad \begin{split}
			\psi^{-1}_{G_2}\circ h_{\der{G},1}=&h'_1\\=&g_1\circ i_0\end{split} \]
		yielding the thesis. \qedhere 
	\end{enumerate}
\end{proof}

@@

	Let $\X$ be the  
	category associated with this order, which by \Cref{rem:iso} is
	$\mathsf{I(\X)}$-adhesive. Consider a rewriting
	system whose set of rules contains the following 
	\[\xymatrix{a & a \ar[r]^{(a,0)} \ar@{>->}[l]_{\id{a}} & 0 & a & a
		\ar[r]^{(a,b)} \ar@{>->}[l]_{\id{a}} & b }\]
	We can then consider the  following derivation
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$.
	 \[
   \xymatrix@C=25pt{a \ar[d]_{(a,c)}&& \ar@{>->}[d]_{k_0}\ar[ll]_{\id{a}}
		    \ar[r]^{(a,0)} & 0 \ar@/^.35cm/[drrr]|(.315)\hole_(.45){\id{0}}
		    \ar[dr]|(.3)\hole_{\id{0}} && a \ar@/_.35cm/[dlll]^(.45){\hspace{5pt}(a,c)}
		     \ar[dl]|(.3)\hole^{(a,0)}& a \ar[d]^{(a,0)}\ar@{>->}[l]_{\id{a}}
		     \ar[rr]^{(a,b)} && b \ar[d]^{(b,0)}\\c &&
		     \ar@{>->}[ll]^{\id{c}} c \ar[rr]_{(c,0)}&& 0 &&
		     \ar@{>->}[ll]^{\id{0}} 0 \ar[rr]_{\id{0}}&& 0}\]

@@

\begin{definition}[Switch] 	Let $(\X, \R)$ be a left-linear rewriting system, with $\X$ an $\mathcal{M}$-adhesive category.
	Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the diagram below, where the cone with vertex $P$ is a pullback and $u_0\colon K_0\to P$ and $u_1\colon K_1\to P$ are the arrows built in \Cref{prop:tec}.
	\[\xymatrix@R=15pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0 \ar @/^.35cm/[ddrr]_(.2){u_0}
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}|(.68)\hole
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}|(.68)\hole 
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar @/_.35cm/[ddll]^(.2){u_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}|(.7)\hole&& G_1 && \ar@{>->}[ll]^{f_1}|(.7)\hole D_1
		\ar[rr]_{g_1}&& G_2 \\ & & & & P \ar[urr]_{p_1} \ar@{>->}[ull]^{p_0}}
	\]

	Let $\der{E}=\{\dder{E}_i\}_{i=0}^1$, depicted below, be a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=15pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

@@

\begin{remark}\label{rem:un}
	Notice that, since $f'_1$ belongs to $\mathcal{M}$, there is at most one arrow $z_1\colon P\to D'_1$ such that $f'{1}\circ z_1=g'_0\circ z_0$.
\end{remark}

We are now going to characterize switches along a given independence pair by means of a property of its colimit.

\begin{lemma}\label{lem:equi}
Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation of length $2$ and $(i_0, i_1)$ an independence pair between $\dder{D}_0$ and $\dder{D}_1$. Given a pre-switch $\der{E}=\{\dder{E}_{i}\}_{i=0}^1$ for $\der{D}$ along $(i_0, i_1)$, the following are equivalent:
	\begin{enumerate}
		\item $\der{E}$ is a switch;
		\item if $(P, \{p_i\}_{i=0}^1)$ is a pullback for $f_{\der{D},1}$ along $g_{\der{D},0}$, then \[\iota_{\der{E}, G_2} \circ g_1\circ p_1=\iota_{\der{E}, G'_1}\circ g'_0\circ z_0\]
	\end{enumerate}
\end{lemma}
\begin{proof}$(1 \Rightarrow 2)$. If $\der{E}$ is a switch then we have
	\begin{align*}
		\iota_{\der{E}, G_2} \circ g_1\circ p_1&=\iota_{\der{E}, G_2} \circ g'_1\circ z_1\\&=\iota_{\der{E}, D'_1}\circ z_1\\&=\iota_{\der{E}, G'_1}\circ f'_1\circ z_1\\&=\iota_{\der{E}, G'_1}\circ g'_0\circ z_0
	\end{align*}

@@

	By \Cref{rem:zero1} the inner square is a pullback, proving that the dotted arrow $z_1\colon P\to D'_1$ exists.  Since, by \Cref{lem:colim} $\iota'_{1, G'_1}$ is in $\mathcal{M}$, we conclude that $f'\circ z_1=g'\circ z_0$ as wanted.
\end{proof}
\iffalse 
We turn now to simmetry of the switchability relation. Surely one desires an analog of \Cref{rem:switch} even for switches: this is assured by the following proposition.

\begin{proposition}\label{prop:pb}
	Let $\der{D}:=\{\dder{D}_{i}\}_{i=0}^1$ be the derivation below, with $(i_0, i_1)$ an independence pair between $\dder{D}_0$  and $\dder{D}_1$.
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]

	If $\der{E}:=\{\dder{E}_{i}\}_{i=0}^1$ is a switch for $\dder{D}$ along $(i_0, i_1)$, then the square below is a pullback.
	\[\xymatrix{P  \ar[r]^-{z_1} \ar[d]_{z_0}& D_{\der{E}, 1} \ar@{>->}[d]^{f_{\der{E},1}} \\ D_{\der{E},0} \ar[r]_{g_{\der{E},0}}& G_{\der{E},1}}\]

	In particular, the arrow $z_0\colon P\to D_{\der{E},0}$ belongs to $\mathcal{M}$.
\end{proposition}
\begin{proof}
	Take two arrows $t_0\colon T\to D_{\der{E}_1}$ and $t_1\colon T\to D_{\der{E},0}$ such that \[f_{\der{E},1}\circ t_1=g_{\der{E},0}\circ t_0\]

	Thus we have
	\begin{align*}
		\iota_{\der{E}, G_0}\circ f_{\der{E},0}\circ t_0=&\iota_{\der{E}, D_{\der{E},0}}\circ t_0\\=&\iota_{\der{E}, G_1}\circ g_{\der{E},0}\circ t_0\\=&\iota_{\der{E}, G_1}\circ f_{\der{E},1}\circ t_1\\&= \iota_{\der{E}, D_{\der{E},1}}\circ t_1\\&=\iota_{\der{E}, G_2}\circ g_{\der{E},1} \circ t_1
	\end{align*}

	By \Cref{cor:ele} we know that there exists a unique $t\colon T\to D_{}$
\end{proof}

\fi

@@

\begin{proposition}\label{prop:pretosw} 	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose
	that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$ and
	$\der{F}=\{\dder{F}_i\}_{i=0}^1$ are pre-switches of $\dder{D}_0$ and
	$\dder{D}_1$ along the same independence pair $(i_0, i_1)$. Let $\{\phi_X\}_{X\in \Deltamin(\der{E})}$ be the abstraction equivalence built in \Cref{thm:switch_uni}. If $\der{F}$ is a switch then also $\der{E}*\phi_{G_2}$ is one.
\end{proposition}
\begin{proof}
	By \Cref{rem:absequi} and \Cref{lem:abs} we know that $\der{E}*\phi_{G_2}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$. Let $(\tpro{E}, \{\iota_{\der{E}, X}\}_{X\in \Deltamin(\der{E})})$, $(\tpro{F}, \{\iota_{\der{F}, X}\}_{X\in \Deltamin(\der{F})})$ and $(\lpro \der{E}*\phi_{G_2}\rpro , \{\iota'_{X}\}_{X\in \Deltamin(\der{E}*\phi_{G_2})})$ be colimiting cocones. By \Cref{rem:absequi,rem:abscol} we have isomorphisms $\Psi \colon \tpro{E}\to \lpro \der{E}*\phi_{G_2}\rpro $ and $\Phi\colon \tpro{E}\to \tpro{F}$ such that 
	\[\begin{split}
		\Psi\circ \iota_{\der{E}, G_0}&=\iota'_{G_0}\\\Phi\circ \iota_{\der{E}, G_0}&=\iota_{\der{F}, G_0}
	\end{split} \quad \begin{split}
	\Psi \circ \iota_{\der{E}, G_{\der{E}, 1}}&=\iota'_{G_1}\\
	\Phi \circ \iota_{\der{E}, G_{\der{E}, 1}}&=\iota_{\der{F}, G_{\der{F}, 1}}\circ \phi_{G_{\der{E}, 1}}
	\end{split} \quad \begin{split}
	\Psi \circ \iota_{\der{E}, G_2}&=\iota'_{G_2}\circ \phi_{G_2}\\
	\Phi \circ \iota_{\der{E}, G_2}&=\iota_{\der{F}, G_2}\circ \phi_{G_2}
	\end{split}\]

 	Let $(P, \{p_0, p_1\})$ be the pullback of $g_{0}$ along $f_1$, so that by \Cref{lem:cose} 	and the definition of switch we have the diagrams below.

 		\[\xymatrix{P\ar[r]^{z_0} \ar@{>->}[d]_{p_0}& D'_{0} \ar@{>->}[d]^{f'_{0}} & P \ar@{>->}[d]_{p_0}\ar[r]^-{z_{\der{F},0}} & D_{\der{F}, 0}\ar@{>->}[d]^{f_{\der{F},0}} & D_{\der{F},0}  \ar[d]_{g_{\der{F},0}}& P \ar[r]^{p_1} \ar[l]_-{z_{\der{F},0}} \ar[d]^{z_{\der{F},1}} & D_1 \ar[d]^{g_1}\\ D_0\ar@{>->}[r]_{f_0} & G_0 & D_0\ar@{>->}[r]_{f_0} & G_0 & G_{\der{F},2} & D_{\der{F},1} \ar[r]_-{g_{\der{F},1}} \ar@{>->}[l]^-{f_{\der{F},1}} & G_2}\]

 	Now, by \Cref{thm:switch_uni} we know that $f'_{0}=f_{\der{F},0}\circ \phi_{D_{\der{E},0}}$, so that 
 	\begin{align*}
 		f_{\der{F},0}\circ \phi_{D_{\der{E},0}}\circ z_0&=f'_{0}\circ z_0\\&=f_0\circ p_0=f_{\der{F},0}\circ z_{\der{F},0}
 	\end{align*}
 	and thus $z_{\der{F},0}=\phi_{D_{\der{E},0}}\circ z_0$. Since $\der{F}$ is a switch we can use \Cref{lem:equi} to get:
	\begin{align*}
		\iota'_{G_2}\circ g_1\circ p_1=&\Psi\circ \iota_{\der{E}, G_2}\circ \phi^{-1}_{G_2}\circ g_1\circ p_1\\=&\Psi\circ \Phi^{-1}\circ \iota_{\der{F}, G_2}\circ g_1\circ p_1\\=&\Psi \circ \Phi^{-1}\circ \iota_{\der{F}, G_{\der{F}, 1}}\circ g_{\der{F}, 0}\circ z_{\der{F},0}\\=&\Psi \circ \iota_{\der{E}, G_{\der{E}, 1}}\circ \phi^{-1}_{G_{\der{E}, 1}}\circ g_{\der{F}, 0}\circ z_{\der{F},0}\\=&\Psi \circ \iota_{\der{E}, G_{\der{E}, 1}}\circ g_{\der{E},0}\circ \phi^{-1}_{D_{\der{E}, 0}}\circ  z_{\der{F},0}\\=&\iota'_{G_1}\circ g'_{0} \circ \phi^{-1}_{D_{\der{E}, 0}} \circ z_{\der{F},0}\\=&	\iota'_{G_1}\circ g'_{0} \circ z_{0}\end{align*}

	The thesis now follows using again from \Cref{lem:equi}.
	\end{proof}

@@

@@

\begin{corollary}\label{cor:refl}
	Let $\der{D}:=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and let $(i_0, i_1)$ be a strong independence pair between $\dder{D}_0$ and $\dder{D}_1$. Suppose that the derivation $\der{E}$ below is a switch for $\der{D}$ along $(i_0, i_1)$.	
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]
